---
title: "Introduction to R"
author: "Nick Bartelo"
date: "2/14/2021"
output: html_document
---

## Q1

1. Use the in-built object state.x77. Find out:

* What type of object you’re dealing with? (Name one of the structures explained in Chapter 3) 
* How to peek into it without printing its full content. 
* How to assess the vector type (e.g. string, integer, …)?

We begin by using the shortcut for help, '?', which opens up the website page http://127.0.0.1:10014/library/datasets/html/state.html which contains information about the data. This page gives us information about the columns and other details regarding the matrix.

```{r include=FALSE}
?state.x77
```

The code below tells us what type of object state.x77 is. We see that it is a matrix.

```{r}
class(state.x77)
```

The code below shows us the first few rows of the state.x77 matrix.

```{r}
head(state.x77)
```

The code below tells us what data type the columns are. The output is double which is just double precision floating point numbers.

```{r}
typeof(state.x77)
```


## Q2

2. How can you find out the default values that a function may assume and that may return unexpected results? Use the read.csv function as an example to highlight two default settings you may have to tweak.

To find out the default values a function may assume, you use the following code, with read.csv being the function.

```{r}
?read.csv
```

## Q3

3. Calculate the sum of all state areas that are greater than 12345 sq miles. Use the built-in state.area vector.

The code below answers the question:

```{r}
state.area
sum(state.area > 12345)
```